Blum axioms and nondeterministic computation of functions

### 2015-12-16 paper.tex
paper.tex: Solve the halting problem under Hopcroft and Ullman's definition.

\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}

@@

>>>>

\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}
\newtheorem{proposition}[definition]{Proposition}

@@

The problem with their definition is that $\phi_w(x)$ is allowed to be defined,
even if some branch of computation does not halt.
This allows us to solve the complement of the halting problem.

\begin{proposition}
    The function partial $f: \Sigma^* \times \Sigma^* \to \{0, 1\}$ defined by
    \begin{equation*}
        f(w, x) = \begin{cases}
            1, & \text{if the machine $w$ does not halt on $x$.} \\
            \text{undefined}, &\text{if $w$ never halts on $x$.}
        \end{cases}
    \end{equation*}
    is not computable,
    but can be computed by a nondeterministic Turing machine under
    Hopcroft and Ullman's definition.
\end{proposition}

\begin{proof}
    Notice that the domain in which $f$ is defined
    is the complement under $\Sigma^*$ of the halting problem.
    Therefore,
    we can pair a machine that computes $f$
    with a machine that recognizes (but not decides) the halting problem
    to actually decide the halting problem,
    as follows:

    Let $M$ be the machine that computes $f$,
    and $U$ be the universal Turing machine
    (that, on input $(w, x)$,
    either returns $\pi_w(x)$ if it exists
    or loops forever if it doesn't).
    The machine $M'$ that will solve the halting problem
    will simulate $M$ and $U$ simultaneously,
    one move for each machine.
    Since the domains of $M$ and $U$ are complementary,
    for any input $(w, x)$, exactly one of them will halt.
    If is $U$, then we accept because $w$ halts on $x$;
    otherwise, if it is $M$, we reject because $w$ never halts on $x$.
    This solves the halting problem --- a contradiction.
    Hence, no Turing machine can compute $f$.

    But $f$ can be computed under Hopcroft and Ullman's definition as follows:
    on input $(w, x)$,
    create two branches of computation.
    On the first, write $1$ on the tape and halt.
    On the second, simulate the universal Turing machine $U$ in the input,
    and if $U$ halts, write $0$ on the tape and also halt.

    If $w$ halts on $x$,
    there will be two halting branches of computation
    that writes different values on the tape,
    so, by definition, the function is not defined on this input.
    If $w$ never halts,
    then only the first branch will halt
    (and with $1$ written on the tape),
    so the function is defined on this input and its value is $1$.

    Thus, this machine can, under Hopcroft and Ullman's definition,
    compute the noncomputable function $f$.
\end{proof}
